### 2017-02-25 IARC_Paper/Paper.tex
Add paper and mission rules

\documentclass{article}

\usepackage{geometry}
\geometry{letterpaper}
\usepackage{doc}
\usepackage{url}
\usepackage{lastpage}
\usepackage{graphicx}
\usepackage{fancyhdr}

\title{IARC Research Paper}
\author{Texas Aerial Robotics}
\date{February 25, 2017}

\pagestyle{fancy}
\renewcommand{\headrulewidth}{0pt}
\fancyhead{}
\cfoot{\thepage\ of \pageref{LastPage}}

\begin{document}
\maketitle
\thispagestyle{empty}

\abstract{
Texas Aerial Robotics will compete in Mission 7 of the International Aerial Robotics Competition (IARC) in 2017. The goal is to direct Roombas (iRobot Create 2's), known as ground robots, using a quadcopter. We use an Intel stereo camera (put name here) for vision, a Nvidia Jetson for vision processing, and a Pixhawk for controlling the quadcopter. 
}

\pagebreak
\tableofcontents

\pagebreak

\section{Introduction}
\label{introduction}

Description of mission with link to rules.

\section{Background, Preliminary, and Related Work}

Describe history, work, or projects that serve as direct contributors to the subject of your research paper

\section{Quadcopter}

Vehicle information

\subsection{Hardware}

Probably have a table for what is what.

\subsection{Controls}

\subsection{Vision}

\subsection{Business}

\section{Safety}

\section{Simulations}

\section{Conclusion}

Hopefully we do well at Georgia Tech! gl hf!

\section{Acknowledgements}

\bibliography{Paper}
\bibliographystyle{unsrt}

\end{document}